---
title: "rmd_submission"
output: html_document
date: "2024-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(forcats)
library(readxl)
library(ggplot2)
library(scales)
library(gridExtra)

# Daten laden
gifts_age <- read_csv("./data/tidytuesday-2024-02-13/gifts_age.csv", show_col_types = FALSE)
gifts_gender <- read_csv("./data/tidytuesday-2024-02-13/gifts_gender.csv", , show_col_types = FALSE)
historical_spending <- read_csv("./data/tidytuesday-2024-02-13/historical_spending.csv", show_col_types = FALSE)
data <- read.csv("./data/Covid-19-data/us_covid19_daily.csv")
```

## Introdcution

During the Covid-19 Pandemic people in the US were severely restricted their social life. The Frequency and experience of meeting friends and family dropped, dating, or just going grocery shopping was heavily impaired. To see how these restrictions impacted the spending behavior of people we take a look the spendings of people on Valentines day to see if the pandemic led to an increase in spendings for others. We will then analyze What gits were the most popular and how age is related to what and how much people buy on Valentines day over the course of 12 years (2010 - 2022).

## How does the Covid-19 Pandemic impact the behaviour on Valentines day

```{r, echo=FALSE}
# Ungültige oder fehlende Datumswerte entfernen
data <- data[!is.na(data$date) & data$date != "", ]

# Datum als Dateiformat konvertieren
data$date <- as.Date(as.character(data$date), format="%Y%m%d")

# Daten nach Datum sortieren
data <- data[order(data$date), ]

# Plot erstellen: Positive Fälle pro Tag
ggplot(data, aes(x = date, y = positive)) +
  geom_line(color = "blue") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") + # Monatliche Achse anzeigen
  labs(title = "COVID-19 Fälle pro Tag in den USA",
       x = "Datum",
       y = "Positive Fälle") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # x-Achse schräg stellen
```

We observe that the number of infected people per day increases rapidly in the year 2020. Let's now have a look at the spending behaviour at this point in time.

```{r, echo=FALSE}
# Plot 1: Prozent der Leute, die feiern
g1 <- ggplot(historical_spending, aes(x = Year, y = PercentCelebrating)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  labs(title = "Prozent der Feiernden",
       x = "Jahr",
       y = "Prozent") +
  theme_minimal()

# Plot 2: Durchschnittlicher Ausgabenbetrag pro Person
g2 <- ggplot(historical_spending, aes(x = Year, y = PerPerson)) +
  geom_line(color = "green", size = 1) +
  geom_point(color = "green") +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  labs(title = "Durchschnittliche Ausgaben pro Person",
       x = "Jahr",
       y = "Betrag in $") +
  theme_minimal()

# Beide Plots nebeneinander anzeigen
grid.arrange(g1, g2, ncol = 2)
```

While we can see a clear downwards trend in the Percentage of people celebrating Valentines day before the pandemic, we can see a significant local peek at the year 2020 in terms of the percentage of people celebrating Valentines day in the US. We can even observe a global peek (in the range fo the collected data) ob the average amount people spent on Valentines day in the year 2020 and a clear upwards trend afterwards, following the trend before the pandemic has started. So even though over less and less people celebrate Valentines day over since 2010, the Covid-19 pandemic influenced the spending behavior and the amount of people celebrating positively.

## Does Age affect the spending behaviour?

Now, we want to investigate the connection between the different Age groups and their spending behavior on Valentines day. Therefore we Plot the percentage of each Age group with how much percent of people from this grup spent money on valentines day. Then we fit a linear Model to this data to see if there is a linear connection between the Age and the amount of people from an age group spending money on Valentines day.

```{r, echo=FALSE}
# Sicherstellen, dass "Age" als Faktor behandelt wird und die Labels korrekt sind
age_labels <- levels(as_factor(gifts_age$Age))
gifts_age$Age <- factor(gifts_age$Age, levels = age_labels)

# Plot: Altersgruppen vs. SpendingCelebrating
ggplot(gifts_age, aes(x = Age, y = SpendingCelebrating, group = 1)) +
  geom_line(aes(color = "Data"), size = 1) + # Linie für die Daten
  geom_smooth(aes(color = "Linear Model"), method = "lm", se = FALSE) + # Lineares Modell hinzufügen
  labs(
    title = "Spending Behavior vs. Age",
    x = "Age Groups",
    y = "Percent Spending Money on Valentine's Day",
    color = "Legend"
  ) +
  scale_x_discrete(labels = age_labels) + # Altersgruppen korrekt anzeigen
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "#999999", size = 0.25),
    panel.grid.minor = element_line(color = "#999999", size = 0.25),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )

```

We can clearly see a downwards trend and the linear line fits the data pretty neat. That means that there is a clear, linear connection between age and spending behaviour. This connection is a nearly perfect negative correlation. The older the people the less they tend to spend money on Valentines day.

## Does Gender affect the spending behaviour?

To further investigate the spending behavior on Valentines day we can look at what percentage of each gender spent money on Valentines day and how this is distributed over different items.

```{r, echo=FALSE}
# Labels für die Kategorien festlegen
category_labels <- c("Candy", "Flowers", "Jewelry", "GreetingCards", "EveningOut", "Clothing", "GiftCards")
gifts_gender_long <- gifts_gender %>%
  pivot_longer(cols = -Gender, names_to = "Category", values_to = "Value")

# Einträge, die keiner der Kategorien entsprechen, entfernen
gifts_gender_long <- gifts_gender_long %>%
  filter(!is.na(Category) & Category %in% category_labels)

# Plot erstellen
ggplot(gifts_gender_long, aes(x = Category, y = Value, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(Value, 1), "%")), position = position_dodge(width = 0.9), vjust = -0.5) +
  scale_x_discrete(labels = category_labels) +
  scale_fill_manual(values = c("Men" = "#4989ff", "Women" = "#fd4848")) +
  labs(title = "Average % Spending by Gender and Category", x = "Category", y = "Average % Spending") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "#c2c2c2"),
    panel.grid.minor = element_line(color = "#787878"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.x = element_text(margin = margin(t = -10))
  )
```

While women predominantly spend their money on candy, clothing, and jewelry, men are more likely to spend on greeting cards, candy, and clothing. Although women outspend men in most item categories, the percentage gap in categories where men spend more is notably larger compared to the gap in categories where women spend more. That mean their are certains Items men like to buy significanlty more than women and those are Greeting Cards and GiftCards. \## How are the Spendings distributed to different items on Valentines day?

It might also be interesting to see which items are the most frequently bought ones on Valentines day. Therefore we plot how much each item contributes to the total spendings on items.

```{r, echo=FALSE}
# Daten für das Kuchendiagramm
spending_categories <- tibble(
  Category = c("Candy", "Flowers", "Jewelry", "Greeting Cards", "Evening Out", "Clothing", "Gift Cards"),
  Percent = c(
    mean(gifts_age$Candy, na.rm = TRUE),
    mean(gifts_age$Flowers, na.rm = TRUE),
    mean(gifts_age$Jewelry, na.rm = TRUE),
    mean(gifts_age$GreetingCards, na.rm = TRUE),
    mean(gifts_age$EveningOut, na.rm = TRUE),
    mean(gifts_age$Clothing, na.rm = TRUE),
    mean(gifts_age$GiftCards, na.rm = TRUE)
  )
)

# Kuchendiagramm
ggplot(spending_categories, aes(x = "", y = Percent, fill = Category)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  labs(
    title = "Percent Spending by Category",
    fill = "Category"
  ) +
  theme_void() +
  theme(legend.position = "right")
```

## Trends of Spending by Category

To see how the above distribution changes over the years we can look at the follwing plot.

```{r, echo=FALSE}
# Daten vorbereiten
historical_spending_labels <- levels(as_factor(historical_spending$Year))
historical_spending$Year <- factor(historical_spending$Year, levels = historical_spending_labels)

only_categories <- historical_spending |> select(-any_of(c("PercentCelebrating", "PerPerson")))
longer <- only_categories |> pivot_longer(!Year, names_to = "Category", values_to = "Percentage")
category_labels <- levels(as_factor(longer$Category))
longer$Category <- factor(longer$Category, levels = category_labels)

# Plot 1: Gestapeltes Balkendiagramm
bar_plot <- ggplot(longer, aes(x = Year, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +
  labs(title = "Average % Spending by Category", 
       x = "Year", 
       y = "Average % Spending") +
  theme_minimal()

# Plot 2: Linienplot mit Trendlinien
line_plot <- ggplot(longer, aes(x = Year, y = Percentage, color = Category, group = Category)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", linetype = "dashed", se = FALSE) +
  labs(title = "Spending Trends by Category", 
       x = "Year", 
       y = "Percentage", 
       color = "Category") +
  theme_minimal()

# Beide Plots nebeneinander anzeigen
grid.arrange(bar_plot, line_plot, ncol = 1)
```

The trend lines correspond to the percentages in the "Average % Spending by Category".

To make the trend lines more visible we exclude the actual data points to recieve the below plot.

```{r, echo=FALSE}
historical_spending <- read_csv("./data/tidytuesday-2024-02-13/historical_spending.csv")

only_categories <- historical_spending |> select(-any_of(c("PercentCelebrating", "PerPerson")))
longer <- only_categories |> pivot_longer(!Year, names_to = "Category", values_to = "Spending")
category_labels <- levels(as_factor(longer$Category))
longer$Category <- factor(longer$Category, levels = category_labels)
ggplot(longer, aes(x = Year, y = Spending, shape = Category, color = Category)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Trends of Spending by Category", x = "Year", y = "Average Spending in $") +
    scale_x_continuous(
    breaks = c(2010, 2012, 2014, 2016, 2018, 2020),   # Festlegung der Positionen der Labels
    labels = c("2010", "2012", "2014", "2016", "2018", "2020")  # Eigene Labels
  ) +
    theme_minimal()
```

Interestingly, the popularity of Jewelery, Clothing, Candy and Gift cards increased significantly while the ones of Greeting Cards and Evenings are somewhat the same. \## Correlation between Age and different Item Purchases

## Correlation between age and item popularity

To see if the spending behavior changes with increasing age looking at certain items we plot the correlation of the Age column with each of the item columns.

```{r, echo=FALSE}
# Erstellen einer Liste mit den Variablen, die wir untersuchen möchten
variables <- c("Candy", "GiftCards", "Flowers", "Jewelry", "GreetingCards", "EveningOut", "Clothing")

# Korrelationen zwischen "Age" und den anderen Variablen berechnen
correlations <- sapply(variables, function(var) {
  cor(as.numeric(as.factor(gifts_age$Age)), gifts_age[[var]], use = "complete.obs")
})

# Erstellen eines Dataframes für die Korrelationswerte
correlation_df <- data.frame(
  Variable = variables,
  Correlation = correlations
)

# Plot der Korrelationen
ggplot(correlation_df, aes(x = Variable, y = Correlation)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.3) +
  labs(
    title = "Korrelation zwischen Age und verschiedenen Items",
    x = "Ausgabenkategorie",
    y = "Korrelationskoeffizient"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

While most items actually decrease in popularity with increasing age, Greeting card become much more popular.
