---
title: "rmd_submission"
output: html_document
date: "2024-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(forcats)
library(readxl)

# Daten laden
gifts_age <- read_csv("./data/tidytuesday-2024-02-13/gifts_age.csv", show_col_types = FALSE)
gifts_gender <- read_csv("./data/tidytuesday-2024-02-13/gifts_gender.csv", , show_col_types = FALSE)
historical_spending <- read_csv("./data/tidytuesday-2024-02-13/historical_spending.csv", show_col_types = FALSE)
```

## Introdcution

in this R-Markdown document we investigate the Valentines day Data-sat from the tidy-tuesday challenge.

## Does Age affect the spending behaviour?

First, we want to investigate the connection bewteen thedifferent Age groups and their spending behaviour on Valentines day. Therefore we Plot how much percent of each Agegroup spend X amount of money on valentines day. Then we fit a linear Model to this data to if there is a linear connection bewteen the Age and the amount of people from this age group spending money on Valentines day.

```{r, echo=FALSE}
# Sicherstellen, dass "Age" als Faktor behandelt wird und die Labels korrekt sind
age_labels <- levels(as_factor(gifts_age$Age))
gifts_age$Age <- factor(gifts_age$Age, levels = age_labels)

# Plot: Altersgruppen vs. SpendingCelebrating
ggplot(gifts_age, aes(x = Age, y = SpendingCelebrating, group = 1)) +
  geom_line(aes(color = "Data"), size = 1) + # Linie für die Daten
  geom_smooth(aes(color = "Linear Model"), method = "lm", se = FALSE) + # Lineares Modell hinzufügen
  labs(
    title = "Spending Behavior vs. Age",
    x = "Age Groups",
    y = "Percent Spending Money on Valentine's Day",
    color = "Legend"
  ) +
  scale_x_discrete(labels = age_labels) + # Altersgruppen korrekt anzeigen
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "#999999", size = 0.25),
    panel.grid.minor = element_line(color = "#999999", size = 0.25),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )
```

## Does Gender affect the spending behaviour?

Next, we wanted to see how the spending behaviour differs between genders.

```{r, echo=FALSE}
# Labels für die Kategorien festlegen
category_labels <- c("Candy", "Flowers", "Jewelry", "GreetingCards", "EveningOut", "Clothing", "GiftCards")
gifts_gender_long <- gifts_gender %>%
  pivot_longer(cols = -Gender, names_to = "Category", values_to = "Value")

# Einträge, die keiner der Kategorien entsprechen, entfernen
gifts_gender_long <- gifts_gender_long %>%
  filter(!is.na(Category) & Category %in% category_labels)

# Plot erstellen
ggplot(gifts_gender_long, aes(x = Category, y = Value, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(Value, 1), "%")), position = position_dodge(width = 0.9), vjust = -0.5) +
  scale_x_discrete(labels = category_labels) +
  scale_fill_manual(values = c("Men" = "#4989ff", "Women" = "#fd4848")) +
  labs(title = "Average % Spending by Gender and Category", x = "Category", y = "Average % Spending") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "#c2c2c2"),
    panel.grid.minor = element_line(color = "#787878"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.x = element_text(margin = margin(t = -10))
  )
```

## How are the Spendings distributed to different items on Valentines day?

It might also be interesting to see which items are the most frequently bought ones on Valentines day. Therefore we plot how much percent each item contributes to the total spendings.

```{r, echo=FALSE}
# Daten für das Kuchendiagramm
spending_categories <- tibble(
  Category = c("Candy", "Flowers", "Jewelry", "Greeting Cards", "Evening Out", "Clothing", "Gift Cards"),
  Percent = c(
    mean(gifts_age$Candy, na.rm = TRUE),
    mean(gifts_age$Flowers, na.rm = TRUE),
    mean(gifts_age$Jewelry, na.rm = TRUE),
    mean(gifts_age$GreetingCards, na.rm = TRUE),
    mean(gifts_age$EveningOut, na.rm = TRUE),
    mean(gifts_age$Clothing, na.rm = TRUE),
    mean(gifts_age$GiftCards, na.rm = TRUE)
  )
)

# Kuchendiagramm
ggplot(spending_categories, aes(x = "", y = Percent, fill = Category)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  labs(
    title = "Percent Spending by Category",
    fill = "Category"
  ) +
  theme_void() +
  theme(legend.position = "right")
```

## Correlation between Age and different Item Purchases

Now lets look at the correlation between the Age Attribute and different items on valentines day:

```{r, echo=FALSE}
# Erstellen einer Liste mit den Variablen, die wir untersuchen möchten
variables <- c("Candy", "GiftCards", "Flowers", "Jewelry", "GreetingCards", "EveningOut", "Clothing")

# Korrelationen zwischen "Age" und den anderen Variablen berechnen
correlations <- sapply(variables, function(var) {
  cor(as.numeric(as.factor(gifts_age$Age)), gifts_age[[var]], use = "complete.obs")
})

# Erstellen eines Dataframes für die Korrelationswerte
correlation_df <- data.frame(
  Variable = variables,
  Correlation = correlations
)

# Plot der Korrelationen
ggplot(correlation_df, aes(x = Variable, y = Correlation)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.3) +
  labs(
    title = "Korrelation zwischen Age und verschiedenen Ausgabenkategorien",
    x = "Ausgabenkategorie",
    y = "Korrelationskoeffizient"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Historical Spending by Category

```{r, echo=FALSE}
historical_spending_labels <- levels(as_factor(historical_spending$Year))
historical_spending$Year <- factor(historical_spending$Year, levels = historical_spending_labels)

only_categories <- historical_spending |> select(-any_of(c("PercentCelebrating", "PerPerson")))
longer <- only_categories |> pivot_longer(!Year, names_to = "Category", values_to = "Percentage")
category_labels <- levels(as_factor(longer$Category))
longer$Category <- factor(longer$Category, levels = category_labels)
ggplot(longer, aes(x = Year, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +
  labs(title = "Average % Spending by Category", x = "Category", y = "Average % Spending") +
  theme_minimal()
```

## Trends of Spending by Category
Here we can see the linear trends of spending for each category.
```{r, echo=FALSE}
historical_spending <- read_csv("./data/tidytuesday-2024-02-13/historical_spending.csv")

only_categories <- historical_spending |> select(-any_of(c("PercentCelebrating", "PerPerson")))
longer <- only_categories |> pivot_longer(!Year, names_to = "Category", values_to = "Spending")
category_labels <- levels(as_factor(longer$Category))
longer$Category <- factor(longer$Category, levels = category_labels)
ggplot(longer, aes(x = Year, y = Spending, shape = Category, color = Category)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Trends of Spending by Category", x = "Year", y = "Average Spending in $") +
    scale_x_continuous(
    breaks = c(2010, 2012, 2014, 2016, 2018, 2020),   # Festlegung der Positionen der Labels
    labels = c("2010", "2012", "2014", "2016", "2018", "2020")  # Eigene Labels
  ) +
    theme_minimal()
```
